---
title: "Analysis of the USJudgeRatings data set"
author: "Adrien Toulouse & Paul-Antoine GIRARD"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 2: Analysis of the USJudgeRatings dataset

This exercise is open. You are asked to use the tools we have seen together to analyze the USJudgeRatings data set. This data set is provided in the package datasets. Your analysis should be reported here and include:

* an introduction
* a general description of the data
* the use of descriptive statistics
* the use of all techniques we have seen together that might be relevant 
* a conclusion

Overall, your analysis, including the graphs and the codes should not exceed 15 pages in pdf.

### Introduction and exploration of the dataset

The USJudgeRatings dataset contains lawyers’ ratings of state judges in the US Superior Court in 1977.

```{r upload data}
data(USJudgeRatings)
head(USJudgeRatings)
```

```{r Introduction}
str(USJudgeRatings)
```

The data is stored in a dataframe.
We are provided with 43 observations and 12 quantitative variables.
```{r}
colnames(USJudgeRatings)
```

The variables are:

* CONT : Number of contacts of lawyer with judge.
* INTG : Judicial integrity
* DMNR : Demeanor
* DILG : Diligence
* CFMG : Case flow managing
* DECI : Prompt decisions
* PREP : Preparation for trial
* FAMI : Familiarity with law
* ORAL : Sound oral rulings
* WRIT : Sound written rulings
* PHYS : Physical ability
* RTEN : Worthy of retention

An observation is the different ratings received by a judge. All the variables (except the variable CONT) are ranged from 0 to 10. 
```{r}
summary(USJudgeRatings)
```

The last variable, RTEN, seems to conclude the analysis. In fact, it says if the lawyers think that the judge is worthy staying in the US Superior Cour or not.

Let's measure the correlations between the 11 first variables and the variable RTEN.


## Descriptive statistics analysis of the dataset

```{r missing values?}
sum(is.na(USJudgeRatings))
```

There are no missing values in the data frame. 

```{r outliers?}
Outvals = boxplot(USJudgeRatings)
```

We observe the presence of outliers for 10 of the 12 variables (with large values for CONT and with low values for the other variables).

```{r mistakes?}
max(USJudgeRatings$CONT)
rownames(USJudgeRatings)[which.max(USJudgeRatings$CONT)]
```

```{r lowest rating}
min(USJudgeRatings$RTEN)
rownames(USJudgeRatings)[which.min(USJudgeRatings$RTEN)]
```

The judge with the lowest rating for worthiness of retention is judge Bracken with a rating of 4.8 

```{r highest rating}
max(USJudgeRatings$RTEN)
rownames(USJudgeRatings)[which.max(USJudgeRatings$RTEN)]
```
The judge with the highest rating for worthiness of retention is judge Rubinow with a rating of 9.2

```{r summary}
summary(USJudgeRatings)
```

We have a symetric distribution for all of the variables since median and mean are always close.

```{r histograms}
par(mfrow=c(2,2))
hist(USJudgeRatings$CONT, main="CONT")
hist(USJudgeRatings$PREP, main="PREP" )
hist(USJudgeRatings$WRIT, main="WRIT")
hist(USJudgeRatings$RTEN, main="RTEN")
```

```{r dispersion}
round(sqrt(diag(var(USJudgeRatings))),2)
```

Regarding the dispersion, we look at the interquartile range (given by the boxplots) and the empirical standard deviation. Overall, the dispersions are not very high (around 1). 
We find that the variables DMNR and RTEN have the largest standard deviation, while the DECI variable has the smallest.

```{r linear relationships between variables 1}
round(cor(USJudgeRatings),2)
```

```{r linear relationships between variables 2}
library(corrplot)
corrplot(cor(USJudgeRatings))
```
All the variables have strong positive correlation two by two except the variable CONT which is not correlated to all the other variables.
The number of contacts of a lawyer with the judge doesn't seem to explain the ratings received by the judge. 

```{r pairs}
pairs(USJudgeRatings)
```


```{r density}
par(mfrow=c(1,2))
hist(USJudgeRatings$CONT, probability= TRUE, main="Histogram of CONT", xlab="CONT")
d = density(USJudgeRatings$CONT, kernel = 'c', bw = 0.3)
lines(d, col="red")

hist(USJudgeRatings$RTEN, probability= TRUE, main="Histogram of RTEN" , xlab="RTEN")
d = density(USJudgeRatings$RTEN, kernel = 'o', bw = 0.3)
lines(d, col="red")
```


```{r ecdf}
par(mfrow=c(1,2))
plot(ecdf(USJudgeRatings$CONT), verticals = TRUE, do.points = FALSE, main = "ECDF CONT")
plot(ecdf(USJudgeRatings$RTEN), verticals = TRUE, do.points = FALSE, main = "ECDF RTEN")
```

```{r QQ plot}
qqnorm(USJudgeRatings$RTEN)
qqline(USJudgeRatings$RTEN)
```

The QQ plots suggests that the RTEN variable is Gaussian.

## Explaining the RTEN variable with a regression model

We will use RTEN as our dependent variable and try to explain it by fitting a regression model. 
We will try to find which of the other 11 variables explain the best our dependant variable and therefore which criterion are the most important for lawyers when evaluating if a judge is fit to stay at the Supreme Court.







